---
title: "A.Sauk_Biol612_Homework3"
author: "Alexandra Sauk"
date: '2020-01-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Part 1 - ANOVAs

##Killarney Lake Data

##Boxplot
```{r}
library("ggplot2")

Kdata<-read.csv("Killarney.csv")

Lake<-Kdata$Lake
Diversity<-Kdata$Diversity
Status<-Kdata$status

par(mar=c(8.5,6.5,4.1,3.1), mgp=c(3,1,0))
ggplot(Kdata, aes(x=Status, y=Diversity, color=Status)) + geom_boxplot() + labs(y="Plankton Species Diversity", x="Lake Status", title = "Fig. 1: Plankton species diversity for 45 lakes in Killarney Wilderness Park") + scale_color_manual(values = c("#ffbb0e", "#009a00", "#007fff"))
```

From the boxplot, it does look like there are differences in plankton diversity between the three lake types, with acid lakes have the least diversity and the neutral, never acid damaged lakes have the greatest diversity. The recovered lakes appear to increases in diversity nearing the levels of diversity seen in the neutral lakes. 

##ANOVA
```{r}
k.fit=aov(Diversity~Status, data = Kdata)
summary(k.fit)
```

The null hypothesis is that there is no difference in the variance of species diversity between the three lake types.

The observed test statistic is F value of 16.41 and the p value is p=5.43e-06. There is a significant difference in species richness amoung the three lake types. 

##Tukey HSD
```{r}
TukeyHSD(aov(Diversity~Status, data = Kdata))
```

Based on the Tukey HSD test, the species diversity is significantly different between the acid lakes and the recovered lakes as well as the acid lakes and the neutral lakes. The neutral and recovered lakes do not differ significantly in terms of their species diversity. 

##Shapiro-Wilk's test
```{r}
shapiro.test(k.fit$residuals)
```

Since p=0.01415, we reject the null hypothesis that the residuals are normally distributed. Thus, the data do not meet the assumption of normality required for the ANOVA.

##Levene's test
```{r}

library("car")
str(Kdata)
leveneTest(Diversity, Status)

```

Since p=0.3773, we do not reject the null hypothesis that the variances in each groups is the same. Thus, the data meets the assumption that the variances in each group is the same. 

##Esophagael Cancer data

##Two factor ANOVA - tobacco and alcohol
```{r}
Edata<-read.csv("esoph.csv")

Age<-Edata$Age
Alcohol<-Edata$Alcohol
Tobacco<-Edata$Tobacco
ncases<-Edata$ncases
ncontrols<-Edata$ncontrols

e.fit1<-aov(ncases~Tobacco*Alcohol)

summary(e.fit1)
```

Based on the results of the two factor ANOVA, neither alcohol consumption or tobacco use on their own or together, signficantly affect on incidences of esophageal cancer as all of the p values are above 0.05.

##Two factor ANOVA - tobacco and age
```{r}
e.fit2<-aov(ncases~Tobacco*Age)

summary(e.fit2)

TukeyHSD(e.fit2)
```

Based on the results of this two factor ANOVA, tobacco use has a signficant affect on incidences of esophageal cancer as p=0.0228 but based on the Tukey test, the effect of tobacco use is only significant when comparing cancer rates between the 0-9g and 30+g per day users (p=0.0348451). Age has a signficant affect overall on incidences of esophageal cancer as p=6.23e-09, with significant between group differences shown below. There is no significant interaction between age and tobacco use. 

Age groups     p-value
45-54-25-34   0.0028395
55-64-25-34   0.0000002
65-74-25-34   0.0001062
45-54-35-44   0.0251086
55-64-35-44   0.0000035
65-74-35-44   0.0013005
75+-55-64    0.0001115
75+-65-74    0.0145751

#Two factor ANOVA - alcohol and age
```{r}
e.fit3<-aov(ncases~Alcohol*Age)

summary(e.fit3)


```
```{r}
interaction.plot(Alcohol, Age, ncases, legend = T, col=c("#808000","#008000","#008080","#000080","#0000FF","#800080")) + title(main = "Fig. 2: Interaction of alcohol consumption and age \n on number of esophageal cancer cases")
```




Based on the results of this two factor ANOVA, alcohol consumption has a signficant affect on incidences of esophageal cancer as p=0.00892 and age has a signficant affect on incidences of esophageal cancer as p=2.02e-09. There is also a significant interaction between age and alcohol asumption with a p=0.03633. ALcohol consumption appears to interact the most with the three ages groups between 45 and 74 years of age. 

#Part 2 - Regression and Correlation

##Question 13

```{r}
mtcars
library("ggplot2")
ggplot(mtcars, aes(x=mtcars$wt, y=mtcars$mpg)) + geom_point(shape = 2, color=("dark green")) + labs(x="Weight (tonnes)", y="Fuel mileage (mpg)", title = "Fig. 3: Fuel milaege of cars versus weight") 



```

Fuel mileage does appear to depend on vehicle weight as there is a negative trend in fuel mileage as weight increases.

##Linear regression
```{r}
car.fit<-glm(mpg~wt, data=mtcars)
summary.lm(car.fit)
```
##Add regression line
```{r}
ggplot(mtcars, aes(x=mtcars$wt, y=mtcars$mpg)) + geom_point(shape = 2, color=("dark green")) + labs(x="Weight (tonnes)", y="Fuel mileage (mpg)", title = "Fig. 4: Fuel milaege of cars versus weight") + geom_abline(intercept = 37.2851, slope = -5.2445, color="blue")


```

## Line equation

y = 37.2851 - 5.2445x

##Significance

There is a significant negative correlation between the fuel mileage and the weight of cars with p=1.294e-10. 

##Question 14

```{r}
mtcars
library("ggplot2")
ggplot(mtcars, aes(x=mtcars$hp, y=mtcars$qsec)) + geom_point(shape = 5, color=("#A2C93A"), size=2) + labs(x="Horsepower", y="Quarter mile time", title = "Fig. 5: Horsepower of cars versus quarter mile time") 



```

Quarter mile time does appear to depend on horsepower as there is a negative trend in quartner mile time as horsepower increases.

##Linear regression
```{r}
car.fit2<-glm(qsec~hp, data=mtcars)
summary.lm(car.fit2)
```

##Add regression line
```{r}
ggplot(mtcars, aes(x=mtcars$hp, y=mtcars$qsec)) + geom_point(shape = 5, color=("#A2C93A"), size=2) + labs(x="Horsepower", y="Quarter mile time", title = "Fig. 6: Horsepower of cars versus quarter mile time")  + geom_abline(intercept = 20.5564, slope = -0.0185, color="blue")

```


## Line equation

y = 20.5564 - 0.0185x

##Quater mile time for a 300 hp car

```{r}
hp<-300
qt = 20.5564 - 0.0185*hp
qt

```

The estimated quater mile time for a car with 300 horsepower is 15.01 seconds. 

##Significance

There is a significant negative correlation between the quarter mile time and the horsepower of cars with p=5.766e-06. 

```{r}
library("gvlma")

gvlma(qsec~hp, data=mtcars)
```

```{r}
hist(car.fit2$residuals)
```

```{r}
plot(fitted(car.fit2),resid(car.fit2))
abline(0,0)
```


The Kurtosis assumption is not satisfied for this data and based on the histogram of the residuals, this may be due to the data not being sufficiently normally distributed. 

##Question 15

The bacteria_genetics.csv includes genome size and gene number for a number of bacteria in the phylum Tenericutes and the order Erysipelotrichales (phylum Firmicutes) from my M.Sc. thesis work. 

```{r}
bac<-read.csv("bacteria_genetics.csv")
Lifestyle<-bac$Lifestyle
library("ggplot2")
ggplot(bac, aes(x=bac$Gene_Count , y=bac$Genome_Size, color=Lifestyle)) + geom_point(shape = 1, size=2) + labs(x="Number of genes", y="Genome Size (Mb)", title = "Fig. 7: Genome size versus gene number \nfor members of the Tenericutes and Erysipelotrichales") + scale_color_manual(values = c("#A2C257","#A4D3EE","#A2627A","#A68064"))


```

```{r}
summary(lm(bac$Genome_Size~bac$Gene_Count))
r=sqrt(0.9471)
r

```

The correlation coefficient is r=0.973. Since p< 2e-16 for the slope, the correlation between genome size and gene number is significant. 

